Changes in Groupings of Geometric Signs as a Model of Social Dynamics Across Aurignacian Periods

### 2025-03-06 endofquarterpaper.qmd
edits

---
title: "Changes in Groupings of Geometric Signs on Paleolithic Mobile Objects as a Model of Social Dynamics Across Aurignacian Time Periods"
author: Sophie Cooper
format: html
echo: false
execute: 
  warning: false
bibliography: references.bib
zotero: "Signbase-study"
link-citations: true
---

@@

euc_mantel <- vegan::mantel(euc_dist, jac_full, permutations = 1000)
p_value_abundance_correlation <- round(euc_mantel$signif, 3)
r_value_abundance_correlation <- round(euc_mantel$statistic, 3)
```

@@

The data set consists of `r nartifacts` artifacts found at `r nsites` sites in `r ncountries` countries, and records `r nsigntypes` geometric sign type found on these artifacts. We excluded signs if they fell under the "other" category, as the wide range of possibilities that could qualify as "other" means that it would be difficult to establish a definitive connection. We also excluded all data from Willendorf (only one sign present), and Riparo di Fontana Nuova, Muralovka, Shanidar Cave, and Hayonim Cave (due to their extreme geographic distance from the other sites). Lastly, data from Grotte De La Princesse Pauline and Šandalja II was excluded, due to being dated much earlier than the rest of the artifacts.

These exclusions resulted in `r nartifacts_clean` artifacts found at `r nsites_clean` sites in `r ncountries_clean` countries, encapsulating `r nsigntypes_clean` sign types. All our tests were run with presence/absence data, in order to keep the data from being affected by outliers like Vogelherd. We ran a Mantel test between an abundance distance matrix and and a Euclidean distance matrix in order to make sure that the exclusion of abundance would not significantly impact our results [@Lycett_2019], and found that it was not quite statistically significant (P = `r p_value_abundance_correlation`, Mantel R = `r r_value_abundance_correlation`).

@@

@@

  x_df <- artifact_data_unique %>% 
    left_join(signbase_full_clean,
              by = "site_name")
  countries <- x_df$group
  unique_countries <- unique(countries)
  color_palette <-  RColorBrewer::brewer.pal(length(unique_countries), "Set3")
  row_colors <- setNames(color_palette, unique_countries)
  result <- concentrate(artifact_matrix, country = countries)
  concentrated_matrix <- result$matrix
  reordered_countries <- result$country
  countries_colors <- row_colors[countries]
  country_colors <- row_colors[reordered_countries]

@@

strength_function <- function(artifact_data, group_data, type = "table"){
  artifact_data <- artifact_data %>% 
    select_if(~ !is.numeric(.) || sum(.) != 0)

  set.seed(500)
  jac <- vegdist(artifact_data, "jaccard", binary = TRUE)
  dm <- as.matrix(jac)

  if(type == "table"){
  modul_matrix <-
    graph_from_adjacency_matrix(1 - dm, mode = "undirected")
  mem <- as_membership(parse_number(group_data$group))
  as_group <- modularity(modul_matrix, membership = mem)

@@

  return(final_table)}

  if(type == "dispersion"){
    set.seed(500)
    dispersion <- betadisper(jac, group = group_data$group)
    disp_plot<- plot(dispersion, hull = FALSE, ellipse=TRUE)
    return(disp_plot)}
}

@@

```{r}
##Modularity test
trans_strength <- strength_function(trans_artifact_data, trans_unique_data, type = "table")
proto_strength <- strength_function(proto_artifact_data, proto_unique_data, type = "table")
early_strength <- strength_function(early_artifact_data, early_unique_data, type = "table")
evolved_strength <- strength_function(evolved_artifact_data, evolved_unique_data, type = "table")

@@

modularity_table <- full_table %>% 
  dplyr::select(`time period`, modularity)

kable(modularity_table)
```

```{r}
##permanova test
permanova_table <- full_table %>% 
  dplyr::select(`time period`, `permanova r`, `permanova p`)

kable(permanova_table)
```

@@

```{r fig.height= 8, fig.width=5}
#| label: fig-perm-disp
#| fig-cap: "Multivariate homogeneity of group dispersions"

trans_strength_plot <- strength_function(trans_artifact_data, trans_unique_data, type = "dispersion")
proto_strength_plot <- strength_function(proto_artifact_data, proto_unique_data, type = "dispersion")
early_strength_plot <- strength_function(early_artifact_data, early_unique_data, type = "dispersion")
evolved_strength_plot <- strength_function(evolved_artifact_data, evolved_unique_data, type = "dispersion")

>>>>

---
title: "Changes in Groupings of Geometric Signs as a Model of Social Dynamics Across Aurignacian Periods"
author: Sophie Cooper
format: html
echo: false
execute: 
  warning: false
bibliography: references.bib
zotero: "Signbase-study"
link-citations: true
---

@@

euc_mantel <- vegan::mantel(euc_dist, jac_full, permutations = 1000)
p_value_abundance_correlation <- round(euc_mantel$signif, 3)
r_value_abundance_correlation <- round(euc_mantel$statistic, 3)

##mantel chronological
chron_df <- signbase_full_clean %>% 
  column_to_rownames("id") %>% 
  dplyr::select(MedianBP) %>% 
  dist( method = "euclidean")

chron_sign_df <- signbase_full_clean %>% 
  column_to_rownames("id") %>% 
  dplyr::select(line:star) %>% 
  vegdist("jaccard")

chron_mantel <- vegan::mantel(chron_df, chron_sign_df, permutations = 1000)

chron_p <- round(chron_mantel$signif, 3)
chron_r <- round(chron_mantel$statistic, 3)

@@

The data set consists of `r nartifacts` artifacts found at `r nsites` sites in `r ncountries` countries, and records `r nsigntypes` geometric sign type found on these artifacts. We excluded signs if they fell under the "other" category, as the wide range of possibilities that could qualify as "other" means that it would be difficult to establish a definitive connection. We also excluded all data from Willendorf (only one sign present), and Riparo di Fontana Nuova, Muralovka, Shanidar Cave, and Hayonim Cave (due to their extreme geographic distance from the other sites). Lastly, data from Grotte De La Princesse Pauline and Šandalja II was excluded, due to being dated much earlier than the rest of the artifacts.These exclusions resulted in `r nartifacts_clean` artifacts found at `r nsites_clean` sites in `r ncountries_clean` countries, encapsulating `r nsigntypes_clean` sign types.

@@

All our tests were run with presence/absence data, in order to keep the data from being affected by outliers like Vogelherd. We ran a Mantel test between an Euclidean abundance distance matrix and a jaccard distance matrix of the sign types in order to make sure that the exclusion of abundance would not significantly impact our results [@Lycett_2019], and found that it was not quite statistically significant (P = `r p_value_abundance_correlation`, Mantel R = `r r_value_abundance_correlation`). 
Another Mantel test was run between a euclidean distance matrix for MedianBP and the jaccard sign type matrix in order to determine whether there was a definitive correlation between distance in time and sign types. The Mantel test found that there was a statistically significant correlation (p = `r chron_p`, Mantel-R = `r chron_r`).

@@

  countries <- artifact_data_unique$group
  unique_countries <- unique(countries)
  color_palette <-  RColorBrewer::brewer.pal(length(unique_countries), "Set3")
  row_colors <- setNames(color_palette, unique_countries)
  result <- concentrate(artifact_matrix, country = countries)
  concentrated_matrix <- result$matrix
  reordered_countries <- result$country
  countries_colors <- row_colors[countries]
  country_colors <- row_colors[reordered_countries]

@@

strength_function <- function(artifact_data, group_data){
  artifact_data <- artifact_data %>% 
    select_if(~ !is.numeric(.) || sum(.) != 0)

  set.seed(500)
  jac <- vegdist(artifact_data, "jaccard", binary = TRUE)
  dm <- as.matrix(jac)
  modul_matrix <-
    graph_from_adjacency_matrix(1 - dm, mode = "undirected")
  mem <- as_membership(parse_number(group_data$group))
  as_group <- modularity(modul_matrix, membership = mem)

@@

  return(final_table)
}

@@

```{r}
##Modularity test
trans_strength <- strength_function(trans_artifact_data, trans_unique_data)
proto_strength <- strength_function(proto_artifact_data, proto_unique_data)
early_strength <- strength_function(early_artifact_data, early_unique_data)
evolved_strength <- strength_function(evolved_artifact_data, evolved_unique_data)

@@

strength_table <- full_table %>% 
  dplyr::select(`time period`, modularity, `permanova r`, `permanova p`)

kable(full_table)
```

@@
